https://www2.stat.duke.edu/\~fl35/teaching/610-23F/docs/project/case-study-II.html

# Load

```{r, include=FALSE}
library(tidyverse)
library(lme4)
library(pps)

vote = read.csv('voter_stats_20201103.txt', sep = "\t")
history = read.csv('history_stats_20201103.txt', sep = "\t")
```

# Cleaning

## Aggregation:

```{r}

# history is actual voters
history = history |> 
  group_by(county_desc,
           voted_party_cd,
           race_code,
           ethnic_code,
           sex_code,
           age) |> 
  summarize(actual_voters = sum(total_voters)) |>
  rename(party_cd = voted_party_cd)

# registered voters
vote = vote |> 
  group_by(
    county_desc,
    party_cd,
    race_code,
    ethnic_code,
    sex_code,
    age
  ) |> 
  summarize(reg_voters = sum(total_voters))
```

## Combine datasets

Notes:

-   some proprtion \> 1, should remove
-   no invalid ages
-   party_cds from here: https://www.fec.gov/campaign-finance-data/party-code-descriptions/
    -   had consitutional and green party not listed in data dict
-   race code P from: https://www.dcf.ks.gov/services/PPS/Documents/PPM_Forms/Section_5000_Forms/PPS5460_Instr.pdf
    -   two missing race code turned into undesignated

```{r}
total_voting = left_join(
  vote,
  history,
  by = c("county_desc", "party_cd", "race_code", "ethnic_code", 
         "sex_code", "age")
  ) |> 
  mutate(proportion = actual_voters / reg_voters, 
         party = case_when(party_cd == "CST" ~ 'constitution',
                           party_cd == "DEM" ~ 'democrat',
                           party_cd == "GRE" ~ 'green',
                           party_cd == "LIB" ~ 'libertarian',
                           party_cd == "REP" ~ 'republican',
                           party_cd == "UNA" ~ 'unaffiliated'),
         race = case_when(race_code == "A" ~ "asian",
                          race_code == "B" ~ "black",
                          race_code == "I" ~ "indian_america",
                          race_code == "M" ~ "two_or_more",
                          race_code == "O" ~ "other",
                          race_code == "W" ~ "white",
                          race_code == "P" ~ "pacific_islander",
                          TRUE ~ "undesignated"),
         ethnicity = case_when(ethnic_code == "HL" ~ 'hispanic_latino',
                               ethnic_code == "NL" ~ 'not_hispanic_latino',
                               ethnic_code == "UN" ~ 'undesignated')) |> 
  ungroup() |> 
  select(-c(race_code, party_cd, ethnic_code))
```

# Sample the counties

-   pps sample based on total registered voters
-   100 total counties in dataset and confirmed online.

```{r}
set.seed(1782)

# first aggregate everything so we can pps sample 
aggregated = total_voting |> 
  group_by(county_desc) |> 
  summarize(total_registered = sum(reg_voters)) |> 
  ungroup()

# draw 37 samples with replacement
county_index = ppss(aggregated$total_registered, 37)

# print county list 
samp_counties = unique(aggregated[county_index, ]$county_desc)
print(samp_counties)
```

# Plots

p. 4 of https://democracync.org/wp-content/uploads/2017/05/WhoVoted2016.pdf

# Resources

https://gvpt.umd.edu/sites/gvpt.umd.edu/files/pubs/Herrnson%20Hanmer%20Niemi%20AJPS%20Ballot%20and%20Errors.pdf
